---
title: "Project1R"
author: "Shunichi Araki, Nicholas David Gabriele, David Raj, Jonathan Simon Wagner"
output: html_document
---

Load data and perform basic cleaning tasks:

```{r}
setwd("C:/Users/sirsl/OneDrive - National University of Singapore/NUS MSBA/Semester 1/DSC5101/Assignments/Econ Assignment 1/")
data.df <- read.csv('Project1Data.csv')
data.df <- subset(data.df, data.df$mean_net_ticket_price < 4000)  # Exclude unrealistic outliers
data.df$num_bookings <-  log(data.df$num_bookings)  # Perform log transform of quantity
data.df$mean_net_ticket_price <-  log(data.df$mean_net_ticket_price)  # Perform log transform of price
data.df <- data.df[order(-data.df[, 3]), ]  # Sort by decreasing mean ticket price

head(data.df)  # Display first 6 rows
```

Create the main reduced form model:

```{r}
price.reduced.form <- lm(mean_net_ticket_price ~ inv, data = data.df)  # Regress IV on endogenous variable
data.df$predicted_price <- predict(price.reduced.form, newx = X)  # Store predicted values to dataframe
data.df$dtd_band <- cut(data.df$dtd, breaks = c(-1, 5, 40, Inf), labels = FALSE)  # Band using specified cuts

summary(price.reduced.form)  # Display summary of model parameters and statistics
```

Create segmented reduced form models:

```{r}
data.df1 <- subset(data.df, data.df$dtd_band == 1)  # Divide dataset into 3 groups by band
data.df2 <- subset(data.df, data.df$dtd_band == 2)
data.df3 <- subset(data.df, data.df$dtd_band == 3)

price.reduced.form.short <- lm(mean_net_ticket_price ~ inv, data.df1)  # Regress IV on endogenous variable
price.reduced.form.medium <- lm(mean_net_ticket_price ~ inv, data.df2)
price.reduced.form.long <- lm(mean_net_ticket_price ~ inv, data.df3)

data.df1$predicted_price <- predict(price.reduced.form.short, newx = X)  # Store predicted values to dataframe
data.df2$predicted_price <- predict(price.reduced.form.medium, newx = X)
data.df3$predicted_price <- predict(price.reduced.form.long, newx = X)

summary(price.reduced.form.short)  # Display summary of model parameters and statistics
summary(price.reduced.form.medium)
summary(price.reduced.form.long)
```

Create main model OLS (for comparison purposes only):

```{r}
structural.form <- lm(num_bookings ~ mean_net_ticket_price, data.df)  # Regress price on quantity

summary(structural.form)  # Display summary of model parameters and statistics
```

Create segmented OLS equations (for comparison purposes only):

```{r}
structural.form.short <- lm(num_bookings ~ mean_net_ticket_price, data.df1)  # Regress price on quantity
structural.form.medium <- lm(num_bookings ~ mean_net_ticket_price, data.df2)
structural.form.long <- lm(num_bookings ~ mean_net_ticket_price, data.df3)

summary(structural.form.short)  # Display summary of model parameters and statistics
summary(structural.form.medium)
summary(structural.form.long)
```

Create main model structural equation:

```{r}
struc.2sls.form <- lm(num_bookings ~ predicted_price, data = data.df)

summary(struc.2sls.form)  # Display summary of model parameters and statistics
```

Create segmented model structural equations:

```{r}
struc.short.2sls.form <- lm(num_bookings ~ predicted_price, data.df1)
struc.medium.2sls.form <- lm(num_bookings ~ predicted_price, data.df2)
struc.long.2sls.form <- lm(num_bookings ~ predicted_price, data.df3)

summary(struc.short.2sls.form)  # Display summary of model parameters and statistics
summary(struc.medium.2sls.form)
summary(struc.long.2sls.form)
```

Create 2SLS main model from built-in R packages with hypothesis tests:

```{r}
library(AER)
library(zoo)

direct.2sls.form <- ivreg(num_bookings ~ mean_net_ticket_price | inv, data = data.df)

summary(direct.2sls.form, diagnostics = TRUE)  # Display summary of model parameters and statistics
```

Create 2SLS segmented models from built-in R packages with hypothesis tests:

```{r}
direct.short.2sls.form = ivreg(num_bookings ~ mean_net_ticket_price | inv, data = data.df1)
direct.medium.2sls.form = ivreg(num_bookings ~ mean_net_ticket_price | inv, data = data.df2)
direct.long.2sls.form = ivreg(num_bookings ~ mean_net_ticket_price | inv, data = data.df3)

summary(direct.short.2sls.form, diagnostics = TRUE)  # Display summary of model parameters and statistics
summary(direct.medium.2sls.form, diagnostics = TRUE)
summary(direct.long.2sls.form, diagnostics = TRUE)
```

Understand covariance between variables using data visualization:

```{r}
library(ggplot2)
library(tidyr)

data.df[c(1:4)] %>%  # Look at each variable's distribution versus target
  gather(-mean_net_ticket_price, key = "var", value = "value") %>%
  ggplot(aes(x = value, y = mean_net_ticket_price, color = mean_net_ticket_price)) +
	geom_point() +
  scale_y_log10() +  # Apply log transformation
	facet_wrap(~ var, scales = "free") +
	theme_bw()
```

Examine distribution of price among chosen segments:

```{r}
data.df$dtd_band <- cut(data.df$dtd, breaks = c(-1, 5, 40, Inf))  # Banding using specified cuts

# Box plot of variable by class
ggplot(data = data.df, aes(x = dtd_band, y = mean_net_ticket_price, fill = dtd_band)) +
  geom_boxplot(alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
